---
title: "Introduction to Package"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Understanding Quality of Life Among Asian Americans: A Case Study Using `Stat108Project2`

## Introduction

Asian Americans are the fastest-growing minority group in the United States, with a phenomenal growth rate of 42.9% between 2000 and 2010, compared to the U.S. total population growth of only 9.3%. According to the U.S. Census Bureau (2012), Asian Americans make up 5.6% of the total U.S. population and are projected to reach 10% by 2050. This rapid growth, combined with the unique challenges faced by new immigrant groups, highlights the need to better understand the social and health needs of Asian Americans.

The `asian_american` dataset in this package provides valuable insights into these issues. It includes detailed demographic, household, employment, emotional well-being, and social/community resource information. Using this dataset, researchers can explore critical questions about quality of life, identify disparities, and inform public health and social policies.

This vignette showcases how the `asianamsurvey` package simplifies the analysis of survey data using this dataset. We'll explore its tools for cleaning, summarizing, and visualizing survey data, with examples drawn from the `asian_american` dataset.

---

## Case Study: Exploring Income and Quality of Life

### Problem Statement

Our goal is to understand how demographic factors like education and English proficiency influence income and quality of life among Asian Americans. Specifically, we'll explore:
- How income varies across education levels.
- Whether English proficiency affects quality-of-life metrics.

---

## Step 1: Data Preparation and Cleaning

### Inspecting the Dataset

First, let's load the dataset and inspect its structure.

```{r}
library(asianamsurvey)

# Load the dataset
data(asian_american)

# View the first few rows
head(asian_american)
```

The dataset includes variables like education, income, english_speaking, and quality_of_life. Many of these variables are categorical and need to be converted for analysis.

### Converting Survey Responses to Numeric
The surveylevels function is helpful for converting categorical responses into numeric scales. For example, we can transform English proficiency ratings into numeric values.

```{r}
# Convert English proficiency levels to numeric
response_scale <- c("Not at all", "Not well", "Well", "Very well")

asian_american <- surveylevels(
  asian_american,
  column = "english_speaking",
  ordered_levels = response_scale
)
```


## Step 2: Summarizing Survey Data
### Summarizing Education Levels and Income
Using the proptable function, we can calculate the proportion of respondents in each education category.

```{r}
# Proportion table for education and income
education_income_table <- proptable(
  asian_american,
  x = "education",
  y = "income",
  type = "proportion"
)

# View the summary
print(education_income_table)
```

## Step 3: Converting variables to numeric
### Converting binary survey variables with inconsistent results to numeric
Using the `makenumeric` function, we can transform binary survey variables into numeric format to facilitate calculations, such as determining the proportion of respondents living in various household arrangements.

In this survey, the options for how the respondent is living are in binary variables, for example, in "spouse", responses are either NA, 0 if not living with spouse, or "Living with spouse" if they are. Most of these are character class, but some are numeric, like "grandparents" or "other relative", so we want to convert them all to binary numeric variables for easy comparison. The `makenumeric` function allows you to specify the variable you want to convert and define the value to treat as `TRUE` (e.g., "Living with spouse"), which is then assigned a numeric value of your choosing (e.g., `1`). You can also specify how to handle other responses (e.g., assign `0` to `FALSE` values) and how to handle `NA` values, which are defaulted to remain the same. This allows for choice, such as if you were handling a different set of data where you want a different true value, like "Dead" in describing mouse weight, to equal "0" instead of "1".


```{r}
# Wrangle data to make living with no one variable binary and numeric
asian_american <- makenumeric(asian_american, "no_one", "Living with no one")

# Wrangle data to make living with spouse variable binary and numeric
asian_american <- makenumeric(asian_american, "spouse", "Living with spouse")

# Wrangle data to make living with children variable binary and numeric
asian_american <- makenumeric(asian_american, "children", "Living with children")

# Wrangle data to make living with grandchildren variable binary and numeric
asian_american <- makenumeric(asian_american, "grand_children", "Living with grandchildren")

# Wrangle data to make living with parent variable binary and numeric
asian_american <- makenumeric(asian_american, "parent", "Living with parents")

# Wrangle data to make living with siblings variable binary and numeric
asian_american <- makenumeric(asian_american, "siblings", "Living with brother/sister")

# Wrangle data to make living with friends variable binary and numeric
asian_american <- makenumeric(asian_american, "friends", "Living with other friends/roommates")
```

## Step 4: Data Visualization

The `scatter` function provides flexible options for visualizing relationships in the survey data. Let's explore some key relationships using different visualization approaches.

### Income and Education Relationships

First, let's examine how income varies with age across different education levels:

```{r fig.width=8, fig.height=6}
scatter(asian_american,
        x = "age",
        y = "income",
        color = "education",
        title = "Income by Age and Education Level",
        xlab = "Age (years)",
        ylab = "Annual Income ($)",
        add_trendline = TRUE)
```

This visualization reveals patterns in income distribution across age groups and education levels, helping identify potential disparities or trends.

### Quality of Life Metrics

We can also explore how quality of life relates to duration of residency and English proficiency:

```{r fig.width=8, fig.height=6}
scatter(asian_american,
        x = "duration_of_residency",
        y = "quality_of_life",
        color = "english_speaking",
        title = "Quality of Life by Duration of Residency",
        xlab = "Years in US",
        ylab = "Quality of Life Score",
        add_trendline = TRUE)
```

### Faceted Analysis of Well-being Indicators

For a more detailed view, we can create faceted plots to examine multiple aspects of well-being:

```{r fig.width=10, fig.height=8}
scatter(asian_american,
        x = "age",
        y = "life_satisfaction",
        color = "english_speaking",
        facet_by = "education",
        title = "Life Satisfaction Across Education Levels",
        xlab = "Age",
        ylab = "Life Satisfaction Score",
        theme_style = "light")
```

### Customizing Visualizations

The scatter function offers various customization options:

- `add_trendline`: Add smoothed trend lines with confidence intervals
- `facet_by`: Create separate plots for different groups
- `theme_style`: Choose between "light" and "dark" themes
- `point_alpha`: Adjust point transparency for better visibility of overlapping points
- `point_size`: Modify point sizes

Here's an example using these features:

```{r fig.width=8, fig.height=6}
scatter(asian_american,
        x = "income",
        y = "achieving_ends_meet",
        color = "education",
        title = "Financial Security by Income Level",
        xlab = "Income",
        ylab = "Ability to Make Ends Meet",
        point_alpha = 0.5,
        point_size = 3,
        add_trendline = TRUE,
        theme_style = "dark")
```

These visualizations help identify patterns and relationships in the survey data, making it easier to:

1. Identify disparities across different demographic groups
2. Understand the relationship between socioeconomic factors and quality of life
3. Explore how language proficiency relates to various outcomes
4. Examine the intersection of multiple factors affecting well-being
