---
title: "Introduction to Package"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Understanding Quality of Life Among Asian Americans: A Case Study Using `Stat108Project2`

## Introduction

Asian Americans are the fastest-growing minority group in the United States, with a phenomenal growth rate of 42.9% between 2000 and 2010, compared to the U.S. total population growth of only 9.3%. According to the U.S. Census Bureau (2012), Asian Americans make up 5.6% of the total U.S. population and are projected to reach 10% by 2050. This rapid growth, combined with the unique challenges faced by new immigrant groups, highlights the need to better understand the social and health needs of Asian Americans.

The `asian_american` dataset in this package provides valuable insights into these issues. It includes detailed demographic, household, employment, emotional well-being, and social/community resource information. Using this dataset, researchers can explore critical questions about quality of life, identify disparities, and inform public health and social policies.

This vignette showcases how the `Stat108Project2` package simplifies the analysis of survey data using this dataset. We'll explore its tools for cleaning, summarizing, and visualizing survey data, with examples drawn from the `asian_american` dataset.

---

## Case Study: Exploring Income and Quality of Life

### Problem Statement

Our goal is to understand how demographic factors like education and English proficiency influence income and quality of life among Asian Americans. Specifically, we'll explore:
- How income varies across education levels.
- Whether English proficiency affects quality-of-life metrics.

---

## Step 1: Data Preparation and Cleaning

### Inspecting the Dataset

First, let's load the dataset and inspect its structure.

```{r}
library(Stat108Project2)

# Load the dataset
data(asian_american)

# View the first few rows
head(asian_american)
```

The dataset includes variables like education, income, english_speaking, and quality_of_life. Many of these variables are categorical and need to be converted for analysis.

### Converting Survey Responses to Numeric
The surveylevels function is helpful for converting categorical responses into numeric scales. For example, we can transform English proficiency ratings into numeric values.

```{r}
# Convert English proficiency levels to numeric
response_scale <- c("Not at all", "Not well", "Well", "Very well")

asian_american <- surveylevels(
  asian_american,
  column = "english_speaking",
  ordered_levels = response_scale
)
```


## Step 2: Summarizing Survey Data
### Summarizing Education Levels and Income
Using the proptable function, we can calculate the proportion of respondents in each education category.

```{r}
# Proportion table for education and income
education_income_table <- proptable(
  asian_american,
  x = "education",
  y = "income",
  type = "proportion"
)

# View the summary
print(education_income_table)
```
